Competing Bandits: The Perils of Exploration under Competition

### 2018-09-05 paper/content/barriers.tex
Half-way done with a pass.

In the temporary monopoly setting, the incumbent firm enjoys a first
mover advantage, which allows the firm to acquire data on potentially
suboptimal actions without incurring an immediate reputational
consequence, and to build up a high reputation before the entrant
enters the market. Thus, the early entry gives the incumbent both a
\textit{data} advantage and a \textit{reputational} advantage over the
entrant. This provides a quantitative perspective on the classic
``first mover advantage" \cite{kerin1992first} in the digital economy
where data can serve as a barrier to entry in many markets.

While the two advantages together serve as strong barriers to entry, a
natural question to ask is whether either one of the advantages alone
suffices as a strong barrier to entry. If both of them are barriers of
entry, which of the two is the stronger?

To answer these questions, we isolate the effect from each of the two
advantages and run two additional experiments. In the first experiment
(called \emph{reputation erased experiment}), the reputation of the
incumbent is artificially reset upon the entry of the entrant {so that
  the incumbent's reputation score is only based on the rewards
  generated in the warm start period}. In the second experiment
(called \emph{information erased experiment}), the information gained
by the incumbent is erased upon the entry of the entrant in the sense
that its posterior is reset to the initial prior. The results of the
experiments are presented in Table~\ref{barrier_exp}, which lead to
our Finding~\ref{barrier-find}.

@@

\begin{finding}\label{barrier-find}
\swedit{During the rounds in which the entrant is present, the incumbent
  with either data or reputational advantage alone gets a larger share
  than the share \gaedit{in the permanent duopoly, regardless of the algorithm she plays}.  When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two
  barriers.}
\OMIT{\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger market share compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}}
\end{finding}

% \swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}

\gaedit{While the data and reputation advantages seem to serve as substitutes for \DEG and \DG, it is interesting that under \TS the data advantage is larger. The reason for this is that, when playing \TS, not only does the early entry allow the incumbent to gather more data than the entrant, but it allows it to gather data on actions \textit{that may be hard to gather data on in competition}. In contrast, purposeful exploration in competition may have immediate adverse reputational consequences. However, in the incumbency period the incumbent faces no immediate reputational consequences from exploration and so has data on a more diversely distributed set of actions but has time to recover lost reputation from selecting the bad arms. As a result, the data advantage is larger for \TS precisely because the advantage that data gives is not just about the quantity of data but also about its distribution across actions.

Why is it that data and reputation alone provide such large barriers to entry? Having more data allows the incumbent to select better actions which allows her to quickly get a high reputation compared to an entrant that both needs to learn and work on building up a reputation. Starting with a large reputation score gives the incumbent a cushion for doing exploration and learning without facing as stark reputational consequences. It is interesting that, for \DEG and \DG, these two effects seem to serve nearly as perfectly substitutes.
}

>>>>

Under temporary monopoly, the incumbent can explore without incurring immediate reputational costs, and build up a high reputation before the entrant appears. Thus, the early entry gives the incumbent both a \textit{data} advantage and a \textit{reputational} advantage over the entrant. While both advantages create a strong barrier to entry, we investigate whether either one is a strong barrier alone, and which one is stronger. Our findings provide a quantitative insight into the classic ``first mover advantage" phenomenon in the digital economy.

For a more succinct terminology, recall that the incumbent enjoys an extended warm start of $X+T_0$ rounds. Call the first $X$ of these rounds the \emph{monopoly period} (and the rest is the proper ``warm start"). The rounds when both firms are competing for customers are called \emph{competition period.}

We run two additional experiments to isolate the effects of the two
advantages mentioned above. The \emph{data-advantage experiment} focuses on the data advantage by, essentially, erasing the reputation advantage. Namely, the data from the monopoly period is not used in the computation of the incumbent's reputation score. Likewise, the \emph{reputation-advantage experiment} erases the data advantage and focuses on the reputation advantage: namely, the incumbent's algorithm `forgets' the data gathered during the monopoly period.

@@

We find that either data or reputational advantage alone gives a substantial boost to the incumbent, compared to permanent duopoly. The results for the Heavy-Tail instance are presented in Table~\ref{barrier_exp}, in the same structure as Table~\ref{tab:ht-incum}. For the other two instances, the results are qualitatively similar.

We can quantitatively define the data (resp., reputation) advantage as the incumbent's market share in the competition period in the data-advantage (resp., reputation advantage) experiment, minus the said share under permanent duopoly, for the same pair of algorithms and the same problem instance. In this language, our findings are as follows.

\begin{finding}\label{barrier-find}
Data advantage and reputation advantage alone are substantially large, across all algorithms and all MAB instances. The data advantage is larger than the reputation advantage when the incumbent chooses \TS; the two advantages are similar when the incumbent chooses \DEG or \DG.
\end{finding}

\ascomment{stopped here.}

\OMIT{ %%%%%
\begin{finding}\label{barrier-find}
The incumbent with either data or reputational advantage alone gets a substantially larger share of agents in the competition period, compared to the said share under permanent duopoly. This holds across all algorithms and all MAB instances.  When the incumbent chooses $\TS$, the data advantage serves as a larger barrier to entry compared to the reputation advantage, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.
\OMIT{\textit{Maintaining either the data or reputation advantage alone still allows the incumbent \gaedit{to retain a larger market share compared to the share they get under permanent duopoly}.\\ \indent When the incumbent plays $\TS$, the data advantage serves as a larger barrier to entry compared to reputation, but when the incumbent plays $\DEG$ or $\DG$ there is no substantial difference between the two barriers.}}
\end{finding}
} %%%%%
% \swcomment{for the first part of the finding, is it true for any algorithm played by the incumbent?}

Our intuition for why the data advantage is larger  under \TS is that, when playing \TS, not only does the early entry allow the incumbent to gather more data than the entrant, but it allows it to gather data on actions \textit{that may be hard to gather data on in competition}. In contrast, purposeful exploration in competition may have immediate adverse reputational consequences. However, in the incumbency period the incumbent faces no immediate reputational consequences from exploration and so has data on a more diversely distributed set of actions but has time to recover lost reputation from selecting the bad arms. As a result, the data advantage is larger for \TS precisely because the advantage that data gives is not just about the quantity of data but also about its distribution across actions.

Why is it that data and reputation alone provide such large barriers to entry? Having more data allows the incumbent to select better actions which allows her to quickly get a high reputation compared to an entrant that both needs to learn and work on building up a reputation. Starting with a large reputation score gives the incumbent a cushion for doing exploration and learning without facing as stark reputational consequences. It is interesting that, for \DEG and \DG, these two effects seem to serve nearly as perfectly substitutes.
